Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-07-17 surfaces-write-up/surface-write-up.tex
merged

%Note that we can view a relation from Equation
%~\ref{eqn:hirz-relations-psi} as an element of $S$ given
%by $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$.  
Let $\Sigma'\subseteq \Sigma$ be the \todo{change all $S$ to sigma's}
set of $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$ satisfying $c_i \ge -\alpha_i d$,
$\bida_i + \sum_{j = 1}^m \bida_j c_j = 0$, and $\bidb_i + \sum_{j=1}^n \bidb_j c_j
= 0$.  Let $\sigma\in \Sigma'$ correspond to the monomial $y_i \prod_{j=1}^n {y_j}^{c_j}$.  Then we can write $\sigma$ as a sum of $r_j e_j$'s for $j\in \Te$ and $r_{j,k}$ for $r\in \Tp, k\in \Tm$.  Let $V$ be the set of $e_i$ $e_{i,j}$ and $e_{j,i}$ as $j$ ranges that exist (i.e. $e_i$ if the hyperplane from lemma \ref{lem:hirz-generators} intersects $e_i$, and the $e_{i,j}$'s or $e_{j,i}$'s if the hyperplane intersects the corresponding edge of the simplex).  A careful computation deduces that if $e_i$ exists
\[
	l
\]
\[
	\lambda + E
\]

\noindent
where $\lambda\in S$ and
\[
	k
\]

\[
	\deg(E) \le \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in \Te \\
	\max_{j \in \Tm} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in \Tp \\
	\max_{j \in \Tp} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in \Tm \end{cases}.
\]

 $= V_1 \cup V_2 \cup V_3$ where
\begin{align*}
	&	V_1 = \{ \{e_i\} \mbox{ if }
	&	V_2= \{e_{i, j}: j \in \Tm \mbox{ if } i	\in \Tp \}
	&	V_3=  \{e_{j, i}: j \in \Tp \mbox{ if } i \in \Tm \}.
\end{align*}

Let $M$ be the set of monomials of $\beta_j$.  Then for $\mu\in M$, the corresponding lattice point is of the form
\[
	E_\mu + \sum_{k\in \Te} r_k e_k + \sum_{\substack{
	h \in \Tp \\ k \in \Tm}} r_{h,k} e_{h,k}
\]
where $E_\mu$ is in degree
\[
	\deg(F) \le \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in \Te \\
	\max_{j \in \Tm} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in \Tp \\
	\max_{j \in \Tp} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in \Tm \end{cases}.
\]

a $\mathbb{R}_{\ge 0}$-combination of elements in
\begin{align*}
	\{e_j : j \in \{1, 2, 3, 4\} \cap \Te\}
	&\cup \; \{e_{j, k}: k \in \Tm, \; j \in \{1, 2, 3, 4\} \cap \Te\} \\
	&\cup \; \{e_{k, j}: k \in \Tp, \; j \in \{1, 2, 3, 4\} \cap \Tm\}.
\end{align*}

Let $F \in R_D$ correspond to the lattice point
\[
	E + \sum_{k \in \Te} \fr(r_k) e_k + \sum_{\substack{h \in \Tp \\
	k \in \Tm}} \fr(r_{h,k}) e_{h,k}
\]

\noindent
where $\fr(r)$ signifies the fractional part of $r$.  Similarly, let $F_\mu
\in R_D$ correspond to the lattice point
\[
	E_\mu + \sum_{\substack{h \in \Tp \\ k \in \Tm}} \fr(r_{h,k}) e_{h,k}.
\]

\noindent
Then $F - \sum_{\mu \in M} F_\mu = 0$ gives us a
relation in degree at most
\[
	\sigma + \deg(F)
\]

\noindent
which divides $(z_i - \beta_i) \prod_{i = 1}^n y_i^{c_i}$.
We can bound $\deg(F)$ by
\[
	\deg(F) \le \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in \Te \\
	\max_{j \in \Tm} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in \Tp \\
	\max_{j \in \Tp} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in \Tm \end{cases}.
\]

>>>>

%Note that we can view a relation from Equation
%~\ref{eqn:hirz-relations-psi} as an element of $S$ given
%by $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$.  
Let $\Sigma'\subseteq \Sigma$ be the \todo{change all $S$ to sigma's}
set of $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$ satisfying $c_i \ge -\alpha_i d$,
$\bida_i + \sum_{j = 1}^m \bida_j c_j = 0$, and $\bidb_i + \sum_{j=1}^n \bidb_j c_j
= 0$.  
Let $\sigma\in \Sigma'$ correspond to the monomial $y_i \prod_{j=1}^n {y_j}^{c_j}$.  Then we can write $\sigma$ as a sum of $r_j e_j$'s for $j\in \Te$ and $r_{j,k}$ for $r\in \Tp, k\in \Tm$.  
For convenience, let $e_j$ (when it exists) be in degree $d_j$ with $j^{th}$ component $-\alpha_j d_j + \kappa_j$ and let $e_{i,j}$ (when it exists) lie in degrees $d_{i,j}$ with $i^{th}$ component $\kappa_{i,j}'$ and $j^{th}$ component $\kappa_{j,i}''$.
Then a careful computation deduces that 
%if $i\in \Te$, $i\in \Tp$, and $i\in \Tm$, then
\[
	\sigma = \lambda + E
\]
where $\lambda \in \Sigma$ and
\begin{equation}\label{eqn:hirz-E-translation}
	E = \begin{cases}
	\frac{1}{\kappa_i} e_i  	&  \mbox{ if }i\in \Te\\
	\sum_{j\in \Tm} r_{i,j} e_{i,j}: r_{i,j}\in \br, \sum_{j\in \Tm} r_{i,j}\kappa_{i,j}' = 1 & \mbox{ if } i\in \Tp\\
	\sum_{j\in \Tp} r_{j,i} e_{j,i}: r_{j,i} \in \br, \sum_{j\in \Tm} r_{j,i} \kappa_{i,j}'' = 1 	& \mbox{ if } i\in \Tm\\
	\end{cases}.
\end{equation}

\noindent
From this, we can deduce that
\[
	\deg(E) \le \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in \Te \\
	\max_{j \in \Tm} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in \Tp \\
	\max_{j \in \Tp} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in \Tm \end{cases}.
\]
Let $M$ be the set of monomials terms of $\beta_j$.  Then for $\mu\in M$, we can further compute that the lattice point corresponding to $\mu \prod_{j=1}^n {f_i}^{c_i}$ is of the form
\[
	\sigma + E_\mu
\]
where $E_\mu$ is defined similarly to $E$ (in particular, they lie in the same degree).

Let $F\in R_D$ corresponding to the lattice point $E$ and let $F_\mu$ correspond to the lattice point $E_\mu$.  
Then $F - \sum_{\mu \in M} F_\mu = 0$ induces a
relation in degree at most
\[
	\sigma + \deg(F)
\]

\noindent
which divides $(z_i - \beta_i) \prod_{i = 1}^n y_i^{c_i}$.
We can bound $\deg(F)$ (which equals $\deg(F_\mu)$ for all $\mu$) by
\[
	\deg(F) = \deg(E) \le \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in \Te \\
	\max_{j \in \Tm} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in \Tp \\
	\max_{j \in \Tp} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in \Tm \end{cases}.
\]
